---
title: "Project I"
author: "Soohee Jung"
date: "6/11/2021"
output:
  github_document:
    toc: TRUE
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# FUNCTIONS  
## Record-API Functions  

```{r message=FALSE}
library(httr)
library(jsonlite)
library(tidyverse)

# to get the franchise ids and the team names
frurl <- GET("https://records.nhl.com/site/api/franchise")
frtext <- content(frurl, "text", encoding = "UTF-8")
frlist <- fromJSON(frtext, flatten=TRUE)
frlist <- as.data.frame(frlist)
frtbl <- tibble(frlist$data.id, frlist$data.fullName)

# record API function
rcdURL <- function(list,recd,type,id){
  if (missing(recd) & missing(type) & missing(id)){
    rcdurl <- paste0("https://records.nhl.com/site/api/",list)
  }
  else if (missing(type) & missing(id)){
    rcdurl <- paste0("https://records.nhl.com/site/api/",list,"-",recd)
  }
  else {
    if (is.numeric(id)){
      rcdurl <- paste0("https://records.nhl.com/site/api/franchise","-",recd,"?cayenneExp=", type, "=", id)
    }
    else {
      id <- filter(filter(frtbl,frtbl[2]==id)[1])
      rcdurl <- paste0("https://records.nhl.com/site/api/franchise","-",recd,"?cayenneExp=", type, "=", id)
    }
  }
  return(rcdurl)
}

rcddt <- function(list,...){
  rcdNHL <- GET(rcdURL(list,...))
  rcdtext <- content(rcdNHL, "text",encoding = "UTF-8")
  rcdlist <- fromJSON(rcdtext, flatten=TRUE)
  rcdlist <- as.data.frame(rcdlist)
  return(rcdlist)
}
```

## Stat-API function  
```{r}
statURL <- function(list,id){
  if (missing(id)){
    staturl <- paste0("https://statsapi.web.nhl.com/api/v1/",list,"?expand=team.stats")
  }
  else{
    if (is.numeric(id)){
      staturl <- paste0("https://statsapi.web.nhl.com/api/v1/",list,"/",id,"?expand=team.stats")
    }
    else {
      id <- filter(filter(frtbl,frtbl[2]==id)[1])
      staturl <- paste0("https://statsapi.web.nhl.com/api/v1/",list,"/",id,"?expand=team.stats")
    }
  }
  return(staturl)
}

statdt <- function(list,...){
  statNHL <- GET(statURL(list,...))
  stattext <- content(statNHL, "text",encoding = "UTF-8")
  statlist <- fromJSON(stattext, flatten=TRUE)
  statlist <- as.data.frame(statlist)
  return(statlist)
}
```

## Wrapper function to call the functions above  
```{r}
# choose record or stat and then put parameters what we want
wrapfnc <- function(fnc,list,...){
  if (fnc=="record"){
    return(rcddt(list,...))
  }
  else if (fnc=="stat"){
    return(statdt(list,...))
  }
  else cat("choose record or stat!!")
}
```

# FACTORS WHICH INFLUENCE TEAM WINNING  
## Choose two franchises to compare with  
```{r}
# Look for interesting results. 
a <- wrapfnc("record","franchise","team-totals") %>% filter(data.gameTypeId==2) %>% 
  mutate(Winchance=data.wins/data.gamesPlayed) %>% select(data.franchiseId,data.gamesPlayed, data.wins, Winchance) 
a[order(a$Winchance),] %>% filter(data.gamesPlayed >2000)
```
ID=1 had higher chance of win than ID=20 had. Let's find the factors which affect team winning!    

## How skater assists affect winning?  
```{r}
# Filter datasets and create new variables
skid1 <- wrapfnc("record","franchise", "skater-records","franchiseId",1) %>% 
  mutate(AvgAssist=mean(data.assists), AvgPenalty=mean(data.penaltyMinutes)) %>% 
  select(data.franchiseId,data.playerId,data.assists,data.goals,data.penaltyMinutes,data.assists,AvgAssist,AvgPenalty)

skid20 <- wrapfnc("record","franchise", "skater-records","franchiseId",20) %>%
  mutate(AvgAssist=mean(data.assists), AvgPenalty=mean(data.penaltyMinutes)) %>% 
  select(data.franchiseId,data.playerId,data.assists,data.goals,data.penaltyMinutes,data.assists,AvgAssist,AvgPenalty)

# Combine two datasets
cosk <- rbind(skid1,skid20)
cosk$data.franchiseId <- as.character(cosk$data.franchiseId)

library(ggplot2)
# Create a graph to see which franchise has more average assists
sk <- ggplot(cosk, aes(x=data.franchiseId, y=AvgAssist))
sk + geom_col(aes(color=data.franchiseId)) + labs(x="Franchise ID", title="< Average Assists by ID>") + 
  theme(legend.position = "none")
# Create a graph to see the relationship between assists and goals
sk2 <- ggplot(cosk, aes(x=data.assists, y=data.goals))
sk2+geom_jitter(aes(color=data.franchiseId))+labs(x="Assists",y="Goals",color="Franchise ID",title="< Assists and Goals >")
```
We can see Franchise ID=1 had more average assists from the first graph. And the second graph tells us the relationship between assists and goals is linear. So, we can say **more assists leads higher chance of winning!**  

## How skater penalty time affect winning?  
```{r}
# Create a graph to see which franchise has more average penalty minutes
skpt <- ggplot(cosk, aes(x=data.franchiseId, y=AvgPenalty))
skpt + geom_col(aes(color=data.franchiseId)) + 
  labs(x="Franchise ID", title="< Average Penalty minute by ID>") + theme(legend.position = "none")

# Create a graph to see the relationship between penalty minutes and goals
skpt2 <- ggplot(cosk, aes(x=data.penaltyMinutes, y=data.goals))
skpt2 + geom_line(aes(color=data.franchiseId)) + 
  labs(x="Penalty Minutes", y="Goals", color="Franchise ID",title="< Penalty minute and Goals >")
```
What a surprising result!! I thought the penalty minutes definitely affects winning but the graphs tell us no relationship between the penalty minutes and the chance of winning. I could not believe that so searched up and found a website. [ESPN/NHL](https://www.espn.com/nhl/statistics/team/_/stat/major-penalties/league/east). Teams can be higher rank regardless of penalty minutes.  


## How goalie goal against affect to winning?  
A good goals against average will fall between 2.00 and 2.70 for NHL goaltenders. Anything between 2.70-3.00 is considered respectable, while below 2.00 is very exceptional.  
```{r}
#goid3 <- wrapfnc("record","franchise", "goalie-records","franchiseId",3)
#goid5 <- wrapfnc("record","franchise", "goalie-records","franchiseId",5)
#goid35 <- rbind(goid3,goid5) %>% select(data.franchiseId,data.playerId,data.mostGoalsAgainstOneGame)
```