---
title: "Project I"
author: "Soohee Jung"
date: "6/11/2021"
output:
  github_document:
    toc: TRUE
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# FUNCTIONS  
## Record-API Functions  

```{r message=FALSE}
library(httr)
library(jsonlite)
library(tidyverse)

# to get the franchise ids and the team names
frurl <- GET("https://records.nhl.com/site/api/franchise")
frtext <- content(frurl, "text", encoding = "UTF-8")
frlist <- fromJSON(frtext, flatten=TRUE)
frlist <- as.data.frame(frlist)
frtbl <- tibble(frlist$data.id, frlist$data.fullName)

# record API function
rcdURL <- function(list,recd,type,id){
  if (missing(recd) & missing(type) & missing(id)){
    rcdurl <- paste0("https://records.nhl.com/site/api/",list)
  }
  else if (missing(type) & missing(id)){
    rcdurl <- paste0("https://records.nhl.com/site/api/",list,"-",recd)
  }
  else {
    if (is.numeric(id)){
      rcdurl <- paste0("https://records.nhl.com/site/api/franchise","-",recd,"?cayenneExp=", type, "=", id)
    }
    else {
      id <- filter(filter(frtbl,frtbl[2]==id)[1])
      rcdurl <- paste0("https://records.nhl.com/site/api/franchise","-",recd,"?cayenneExp=", type, "=", id)
    }
  }
  return(rcdurl)
}

rcddt <- function(list,...){
  rcdNHL <- GET(rcdURL(list,...))
  rcdtext <- content(rcdNHL, "text",encoding = "UTF-8")
  rcdlist <- fromJSON(rcdtext, flatten=TRUE)
  rcdlist <- as.data.frame(rcdlist)
  return(rcdlist)
}
```

## Stat-API function  
```{r}
statURL <- function(list,id){
  if (missing(id)){
    staturl <- paste0("https://statsapi.web.nhl.com/api/v1/",list,"?expand=team.stats")
  }
  else{
    if (is.numeric(id)){
      staturl <- paste0("https://statsapi.web.nhl.com/api/v1/",list,"/",id,"?expand=team.stats")
    }
    else {
      id <- filter(filter(frtbl,frtbl[2]==id)[1])
      staturl <- paste0("https://statsapi.web.nhl.com/api/v1/",list,"/",id,"?expand=team.stats")
    }
  }
  return(staturl)
}

statdt <- function(list,...){
  statNHL <- GET(statURL(list,...))
  stattext <- content(statNHL, "text",encoding = "UTF-8")
  statlist <- fromJSON(stattext, flatten=TRUE)
  statlist <- as.data.frame(statlist)
  statlist <- unnest(unnest(statlist,cols = c(teams.teamStats)),cols = c(splits))
  return(statlist)
}
```

## Wrapper function to call the functions above  
```{r}
# choose record or stat and then put parameters what we want
wrapfnc <- function(fnc,list,...){
  if (fnc=="record"){
    return(rcddt(list,...))
  }
  else if (fnc=="stat"){
    return(statdt(list,...))
  }
  else cat("choose record or stat!!")
}
```

# FACTORS WHICH INFLUENCE TEAM WINNING  
## Choose two franchises to compare with  
```{r}
# Look for interesting results from 'team-total' dataset. 
a <- wrapfnc("record","franchise","team-totals") %>% filter(data.gameTypeId==2 & data.gamesPlayed >2000) %>% 
  mutate(Winchance=data.wins/data.gamesPlayed) %>% select(data.franchiseId,data.gamesPlayed, data.wins, Winchance)
knitr::kable(a)
```
ID=1 had higher chance of win than ID=20 had. Let's find the factors which affect team winning!    

## How skater assists affect winning?  
```{r}
# Filter datasets and create new variables
skid1 <- wrapfnc("record","franchise", "skater-records","franchiseId",1)
skid20 <- wrapfnc("record","franchise", "skater-records","franchiseId",20)

# Combine two datasets
cosk <- rbind(skid1,skid20)
cosk$data.franchiseId <- as.character(cosk$data.franchiseId)

# Summarise skaters assists records
avgassi <- c(ID.1=mean(skid1$data.assists),ID.20=mean(skid20$data.assists))
knitr::kable(avgassi,col.names = "Avg Assists")
```
We can see Franchise ID=1 had more average assists from the data.  

```{r}
library(ggplot2)
# Create a graph to see the relationship between assists and goals
sk <- ggplot(cosk, aes(x=data.assists, y=data.goals))
sk+geom_jitter(aes(color=data.franchiseId))+labs(x="Assists",y="Goals",color="Franchise ID",title="< Assists and Goals >")
```

The franchise ID=1 had more average assists and higher winning chance than ID=20 did. This graph also tells us the relationship between assists and goals is linear. So, we can say **more assists leads higher chance of winning!**  

## How center position skater's penalty time affect goal?  
```{r}
# Summarise center position skater's penalty minutes records
center <- cosk %>% filter(data.positionCode== "C")
centerpt <- ggplot(center, aes(x=data.penaltyMinutes, y=data.goals))
centerpt + geom_line(aes(color=data.franchiseId)) + geom_smooth() +
  labs(x="Penalty Minutes", y="Goals",color="Franchise ID", title="< center skater Penalty minute and Goals >")
```
The skaters who are in a center position scored more goals as they had more penalty minutes. What?? Interesting!  

```{r}
ttotal <- wrapfnc("record","franchise","team-totals")
w <- ggplot(ttotal,aes(data.penaltyMinutes,data.wins))
g <- ggplot(ttotal,aes(data.penaltyMinutes,data.goalsFor))
g+geom_quantile()+labs(x="Penalty Minutes", y="Goals", title="< Penalty minutes and Goals >")
w+geom_quantile()+labs(x="Penalty Minutes", y="Wins", title="< Penalty minutes and Wins >")
```
  
What a surprising result!! I thought the penalty minutes would affects goals and winnings in negative ways, but the graphs tell us totally opposite story.  
